---
title: "Data Analyses"
author: "May"
date: "11/8/2022"
output: html_document
---
Let's analyze the American Kestrel Dataset!

First, lets clear r's brain and add packages
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
```
Now lets import the data
```{r}
kestrels <- read.csv(here("Data", "AMKE DATA TABLE 1.csv"))
```
Now lets analyze the data
First lets do a summary
```{r}
summary(kestrels)
```
Now lets look at a scatterplot of the percent successful boxes to see if there are any outliers
```{r}
ggplot(kestrels, aes(x = BOX, y = X.Successful)) + geom_point(size =1)+
xlab("Box Number") +
ylab("% Success")
```
Does not look like there are too many outliers

```{r}
ggplot(kestrels, aes(x = BOX, y = Num_Chicks_Banded)) + geom_point(size =1)+
xlab("Box Number") +
ylab("Chicks banded")
```
Looks like there could be an outlier lets look at chicks banded stats

Now lets create a histogram of the percent success
```{r}
ggplot(kestrels, aes(x = X.Successful)) + geom_histogram(bins = 10)+
  xlab("Percent Success")
```
There is a gap around 10%




